
```{r}
library(dplyr)
library(lubridate)
library(tidyr)

library(MatchIt)
library(tableone)
library(glmtoolbox)
library(Hmisc)

```


1. Case and control definition
JOIN: Drug_Exposure tables for ICI and steroids
    Case definition: steroid exposure during the continuous ICI treatment window
                     (±60 day rule, 120-day collapsing window, and follow-up up to 60 days after the last ICI date)
    Control definition: no steroid exposure during the continuous ICI treatment window (same collapsing rule)
    
```{r}

ICI <- read.csv("ICI.CSV")[,-1]

patient <- read.csv("patient.CSV")[,-1]
steroid <- read.csv("steroid.CSV")[,-1]
steroid <- steroid %>%
  filter(ROUTE_SOURCE_VALUE != "외용")


print(unique(ICI$DRUG_CONCEPT_ID))
# 36808878 36809461 41370369 42629082 42922128 46276413
# 46276413	4 ML pembrolizumab 25 MG/ML Injection	Drug	Standard			
# 42629082	20 ML atezolizumab 60 MG/ML Injection	Drug	Standard			
# 36809461	2.4 ML durvalumab 50 MG/ML Injectable Solution	Drug	Standard			
# 42922128	2 ML nivolumab 10 MG/ML Injectable Solution	Drug	Standard			
# 41370369	10 ML nivolumab 10 MG/ML Injectable Solution	Drug	Standard			
# 36808878	10 ML durvalumab 50 MG/ML Injectable Solution


# 1) Create continuous ICI exposure windows per patient (collapse gaps of 120 days or less)
# ICI drug categories
ICI_drug <- ICI %>%
  mutate(
    pembrolizumab = as.integer(DRUG_CONCEPT_ID == 46276413),
    nivolumab = as.integer(DRUG_CONCEPT_ID %in% c(42922128, 41370369)),
    atezolizumab = as.integer(DRUG_CONCEPT_ID == 42629082),
    durvalumab = as.integer(DRUG_CONCEPT_ID %in% c(36809461, 36808878))
  )
# Summarise ICI drug exposure indicators at the patient level
ICI_drug <- ICI_drug %>%
  group_by(PERSON_ID) %>%
  summarise(
    pembrolizumab_use = max(pembrolizumab),
    nivolumab_use = max(nivolumab),
    atezolizumab_use = max(atezolizumab),
    durvalumab_use = max(durvalumab),
    .groups = 'drop'
  )


# Function to collapse exposure periods using a 120-day gap threshold
calculate_date_diff120 <- function(df) {
  df %>%
    mutate(
      DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
      DRUG_EXPOSURE_END_DATE = as.Date(DRUG_EXPOSURE_END_DATE)
    ) %>%
    arrange(PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE) %>%
    group_by(PERSON_ID) %>%
    mutate(
      prev_end_date = lag(DRUG_EXPOSURE_END_DATE),
      overlap = DRUG_EXPOSURE_START_DATE <= prev_end_date,
      gap = as.numeric(difftime(DRUG_EXPOSURE_START_DATE, prev_end_date, units = "days"))
    ) %>%
    mutate(
      continuous_group = cumsum(ifelse(is.na(gap) | overlap | gap <= 120, 0, 1))
    ) %>%
    group_by(PERSON_ID, continuous_group) %>%
    summarise(
      ICI_START_DATE = min(DRUG_EXPOSURE_START_DATE),
      ICI_END_DATE = max(DRUG_EXPOSURE_END_DATE),
      .groups = 'drop'
    ) %>%
    ungroup() %>%
    select(-continuous_group)
}


ICI_continu <- select(ICI, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE)
ICI_continu <- unique(ICI_continu)
ICI_collapsed <- calculate_date_diff120(ICI_continu)
ICI_collapsed <- unique(ICI_collapsed)




# 2) Filter steroid exposures occurring within the continuous ICI window plus 60 day
ICI_steroid <- data.frame()

ICI_collapsed$ICI_START_DATE <- as.Date(ICI_collapsed$ICI_START_DATE)
ICI_collapsed$ICI_END_DATE <- as.Date(ICI_collapsed$ICI_END_DATE)



for (i in 1:nrow(ICI_collapsed)) {
  filtered_steroid <- steroid %>% 
  filter(PERSON_ID == ICI_collapsed$PERSON_ID[i]) %>%
  filter((DRUG_EXPOSURE_START_DATE >= ICI_collapsed$ICI_START_DATE[i] &
         DRUG_EXPOSURE_START_DATE <= ICI_collapsed$ICI_END_DATE[i] + days(60))
         |
        (DRUG_EXPOSURE_END_DATE >= ICI_collapsed$ICI_START_DATE[i] &
         DRUG_EXPOSURE_START_DATE < ICI_collapsed$ICI_START_DATE[i])
         )
filtered_steroid$DRUG_EXPOSURE_START_DATE <- as.Date(filtered_steroid$DRUG_EXPOSURE_START_DATE)
filtered_steroid$DRUG_EXPOSURE_END_DATE <- as.Date(filtered_steroid$DRUG_EXPOSURE_END_DATE)
  
 filtered_steroid <- filtered_steroid %>%
   mutate(DRUG_EXPOSURE_START_DATE = if_else(
     DRUG_EXPOSURE_END_DATE >= ICI_collapsed$ICI_START_DATE[i] &
       DRUG_EXPOSURE_START_DATE < ICI_collapsed$ICI_START_DATE[i],
     ICI_collapsed$ICI_START_DATE[i],
     DRUG_EXPOSURE_START_DATE
   ))
  
  ICI_steroid <- rbind(ICI_steroid, filtered_steroid)
}
ICI_steroid <- unique(ICI_steroid)
ICI_collapsed_total <- ICI_collapsed
ICI_steroid_total <- ICI_steroid
write.csv(ICI_steroid_total, file = "ICI_steroid_total.CSV")




calculate_date_diff30 <- function(df) {
  df %>%
    mutate(
      DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
      DRUG_EXPOSURE_END_DATE = as.Date(DRUG_EXPOSURE_END_DATE)
    ) %>%
    arrange(PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE) %>%
    group_by(PERSON_ID) %>%
    mutate(
      prev_end_date = lag(DRUG_EXPOSURE_END_DATE),
      overlap = DRUG_EXPOSURE_START_DATE <= prev_end_date,
      gap = as.numeric(difftime(DRUG_EXPOSURE_START_DATE, prev_end_date, units = "days"))
    ) %>%
    mutate(
      continuous_group = cumsum(ifelse(is.na(gap) | overlap | gap <= 30, 0, 1))
    ) %>%
    group_by(PERSON_ID, continuous_group) %>%
    summarise(
      Steroid_START_DATE = min(DRUG_EXPOSURE_START_DATE),
      Steroid_END_DATE = max(DRUG_EXPOSURE_END_DATE),
      .groups = 'drop'
    ) %>%
    ungroup() %>%
    select(-continuous_group)
}

ICI_steroid_conti <- select(ICI_steroid, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE)
ICI_steroid_conti <- unique(ICI_steroid_conti)
Steroid_collapsed <- calculate_date_diff30(ICI_steroid_conti)
Steroid_collapsed <- unique(Steroid_collapsed)
Steroid_collapsed_total <- Steroid_collapsed







# 3) (2024-10-02) Identify steroid new use within the continuous ICI window:
#     case = steroid use, control = no steroid use.
### For controls, censor if ICI and steroid are used together after the continuous use window ends.



ICI_steroid <- data.frame()
ICI_collapsed$ICI_START_DATE <- as.Date(ICI_collapsed$ICI_START_DATE)
ICI_collapsed$ICI_END_DATE <- as.Date(ICI_collapsed$ICI_END_DATE)

ICI_collapsed <- ICI_collapsed %>%
  group_by(PERSON_ID) %>% 
  slice_min(ICI_START_DATE) %>%
  ungroup()


ICI_collapsed_check <- ICI_collapsed %>%
  group_by(PERSON_ID) %>%
  mutate(
    ICI_END_DATE = if_else(difftime(ICI_END_DATE, ICI_START_DATE, units = "days") > 730, ICI_START_DATE+730, ICI_END_DATE)
  ) %>%
  ungroup()



for (i in 1:nrow(ICI_collapsed_check)) {
  filtered_steroid <- steroid %>% 
  filter(PERSON_ID == ICI_collapsed_check$PERSON_ID[i]) %>%
  filter(DRUG_EXPOSURE_START_DATE >= ICI_collapsed_check$ICI_START_DATE[i] &
         DRUG_EXPOSURE_START_DATE <= ICI_collapsed_check$ICI_END_DATE[i] + days(60)
         |
        DRUG_EXPOSURE_END_DATE >= ICI_collapsed_check$ICI_START_DATE[i] &
         DRUG_EXPOSURE_START_DATE < ICI_collapsed_check$ICI_START_DATE[i]
         )
filtered_steroid$DRUG_EXPOSURE_START_DATE <- as.Date(filtered_steroid$DRUG_EXPOSURE_START_DATE)
filtered_steroid$DRUG_EXPOSURE_END_DATE <- as.Date(filtered_steroid$DRUG_EXPOSURE_END_DATE)
  
 filtered_steroid <- filtered_steroid %>%
   mutate(DRUG_EXPOSURE_START_DATE = if_else(
     DRUG_EXPOSURE_END_DATE >= ICI_collapsed_check$ICI_START_DATE[i] &
       DRUG_EXPOSURE_START_DATE < ICI_collapsed_check$ICI_START_DATE[i],
     ICI_collapsed_check$ICI_START_DATE[i],
     DRUG_EXPOSURE_START_DATE
   ))
  
  ICI_steroid <- rbind(ICI_steroid, filtered_steroid)
}
ICI_steroid <- unique(ICI_steroid)

calculate_date_diff30 <- function(df) {
  df %>%
    mutate(
      DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
      DRUG_EXPOSURE_END_DATE = as.Date(DRUG_EXPOSURE_END_DATE)
    ) %>%
    arrange(PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE) %>%
    group_by(PERSON_ID) %>%
    mutate(
      prev_end_date = lag(DRUG_EXPOSURE_END_DATE),
      overlap = DRUG_EXPOSURE_START_DATE <= prev_end_date,
      gap = as.numeric(difftime(DRUG_EXPOSURE_START_DATE, prev_end_date, units = "days"))
    ) %>%
    mutate(
      continuous_group = cumsum(ifelse(is.na(gap) | overlap | gap <= 30, 0, 1))
    ) %>%
    group_by(PERSON_ID, continuous_group) %>%
    summarise(
      Steroid_START_DATE = min(DRUG_EXPOSURE_START_DATE),
      Steroid_END_DATE = max(DRUG_EXPOSURE_END_DATE),
      .groups = 'drop'
    ) %>%
    ungroup() %>%
    select(-continuous_group)
}

ICI_steroid_conti <- select(ICI_steroid, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE)
ICI_steroid_conti <- unique(ICI_steroid_conti)
Steroid_collapsed <- calculate_date_diff30(ICI_steroid_conti)
Steroid_collapsed <- unique(Steroid_collapsed)



CASE_patient <- data.frame()
CASE_patient <- select(Steroid_collapsed, PERSON_ID)
CASE_patient <- unique(CASE_patient)

patient <- patient %>%
  mutate(TARGET = ifelse(PERSON_ID %in% CASE_patient$PERSON_ID, 1, 0))



### Retain a single ICI continuous-use interval per patient
### Cases: keep the interval that includes the steroid start date; controls: keep the earliest interval.


ICI_collapsed_check <- left_join(ICI_collapsed_check, select(patient, PERSON_ID, TARGET), by = "PERSON_ID")

ICI_collapsed_filtered <- ICI_collapsed_check %>%
  group_by(PERSON_ID) %>%
  filter(!(TARGET == 0 & row_number() != which.min(as.Date(ICI_START_DATE)))) %>%
  ungroup()

Steroid_collapsed_uniq <- Steroid_collapsed %>%
  group_by(PERSON_ID) %>%
  slice(which.min(as.Date(Steroid_START_DATE)))

ICI_collapsed_filtered <- left_join(ICI_collapsed_filtered, Steroid_collapsed_uniq, by = "PERSON_ID")


ICI_collapsed_filtered <- ICI_collapsed_filtered %>%
  group_by(PERSON_ID) %>%
  filter(
    (TARGET == 0) |
    (TARGET == 1 & ICI_START_DATE <= Steroid_START_DATE)
  ) %>%
  ungroup()





###(2024-10-02) For controls, censor if steroid use resumes after the ICI window by truncating COHORT_END_DATE when Steroid_START_DATE occurs earlier.
ICI_collapsed_total_filtered <- left_join(ICI_collapsed_total, select(patient, PERSON_ID, TARGET), by = "PERSON_ID")

Steroid_collapsed_total_uniq <- Steroid_collapsed_total %>%
  group_by(PERSON_ID) %>%
  slice(which.min(as.Date(Steroid_START_DATE)))

ICI_collapsed_total_filtered <- left_join(ICI_collapsed_total_filtered, Steroid_collapsed_total_uniq, by = "PERSON_ID")

ICI_collapsed_total_filtered <- ICI_collapsed_total_filtered %>%
  group_by(PERSON_ID) %>%
  filter(
    (TARGET == 0) |
    (TARGET == 1 & ICI_START_DATE <= Steroid_START_DATE)
  ) %>%
  ungroup()

Control_censor <- ICI_collapsed_total_filtered %>%
  group_by(PERSON_ID) %>%
  filter(
    (TARGET == 1) & (n() >=2)
  ) %>%
  ungroup()


Control_censor <- left_join(Control_censor, select(ICI_collapsed_filtered, PERSON_ID, TARGET), by = "PERSON_ID")
Control_censor <- Control_censor %>%
  filter(TARGET.x != TARGET.y)
patient_newuser <- left_join(patient, select(Control_censor, PERSON_ID, Steroid_START_DATE), by = "PERSON_ID")
patient_newuser <- unique(patient_newuser)

patient_newuser$COHORT_END_DATE <- as.Date(patient_newuser$COHORT_END_DATE)
patient_newuser$Steroid_START_DATE <- as.Date(patient_newuser$Steroid_START_DATE)
patient_newuser <- patient_newuser %>%
  mutate(COHORT_END_DATE = case_when(!is.na(Steroid_START_DATE) & Steroid_START_DATE < COHORT_END_DATE ~ Steroid_START_DATE,
                                     TRUE ~ COHORT_END_DATE))
write.csv(patient_newuser, file = "patient_newuser.CSV")
patient <- select(patient_newuser, PERSON_ID, COHORT_START_DATE, COHORT_END_DATE, TARGET)
 
#unique(ICI_collapsed_filtered$PERSON_ID)

write.csv(Steroid_collapsed, file = "Steroid_collapsed.CSV")
write.csv(ICI_collapsed_filtered, file = "ICI_collapsed_filtered.CSV")









#4)  4) Derive the ICI use cycle per patient and identify the cycle preceding the earliest steroid start date in the case group

ICI_continu <- ICI_continu %>%
  arrange(PERSON_ID, DRUG_EXPOSURE_START_DATE)

ICI_continu <- ICI_continu %>%
  group_by(PERSON_ID) %>%
  mutate(ICI_use_cycle = n()) %>%
  ungroup()  

patient <- left_join(patient, select(ICI_continu, PERSON_ID, ICI_use_cycle), by = 'PERSON_ID')
patient <- unique(patient)

TARGET_index <- Steroid_collapsed %>%
  group_by(PERSON_ID) %>%
  summarise(Steroid_index_date = min(Steroid_START_DATE)) %>%
  ungroup()

patient <- left_join(patient, TARGET_index, by = 'PERSON_ID')
ICI_continu <- left_join(ICI_continu, TARGET_index, by = 'PERSON_ID')


library(dplyr)

# Adjusted function to find the index of the latest DRUG_EXPOSURE_START_DATE before Steroid_index_date
find_latest_record <- function(df, steroid_date) {
  steroid_date <- as.Date(steroid_date)
  valid_dates <- df %>% filter(DRUG_EXPOSURE_START_DATE <= steroid_date)
  if (nrow(valid_dates) == 0) {
    return(NA)
  }
  latest_date <- max(valid_dates$DRUG_EXPOSURE_START_DATE, na.rm = TRUE)
  latest_index <- which.max(valid_dates$DRUG_EXPOSURE_START_DATE == latest_date)
  return(latest_index) 
}

# Applying the function to the ICI_continu dataset to create the ICI_index_cycle column
ICI_cycle <- ICI_continu %>%
  group_by(PERSON_ID) %>%
  mutate(ICI_index_cycle = find_latest_record(cur_data(), Steroid_index_date)) %>%
  ungroup()


patient <- left_join(patient, select(ICI_cycle, PERSON_ID, ICI_index_cycle), by = 'PERSON_ID')
patient <- unique(patient)




#4) Calculate cumulative ICI exposure days and, for cases, the number of days until the steroid index date
ICI_use_days <- select(ICI, PERSON_ID, DRUG_EXPOSURE_END_DATE)
ICI_use_days <- left_join(ICI_use_days, select(patient, PERSON_ID, COHORT_START_DATE), by = "PERSON_ID")
ICI_use_days <- unique(ICI_use_days)

ICI_use_days <- ICI_use_days %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE),
      DRUG_EXPOSURE_END_DATE = as.Date(DRUG_EXPOSURE_END_DATE)
    ) %>%
    mutate(
      latest_date = max(DRUG_EXPOSURE_END_DATE, na.rm = TRUE),
      ICI_use_days = as.numeric(difftime(latest_date, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

    
patient <- left_join(patient, select(ICI_use_days, PERSON_ID, ICI_use_days), by = 'PERSON_ID')
patient <- unique(patient)
patient <- patient %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE),
      Steroid_index_date = as.Date(Steroid_index_date)
    ) %>%
    mutate(
      Index_ICI_days = as.numeric(difftime(Steroid_index_date, COHORT_START_DATE, units = "days")+1)
    )





#5) 5) Capture concomitant anticancer therapy and its timing
# Chemotherapy

chemo_agent <- read.csv("cancer_agent.CSV")[,-1] 

# TKI_ALK
TKI_ALK <- read.csv("TKI_ALK.CSV")[,-1]


cancer_agent <- rbind(chemo_agent, TKI_ALK)
cancer_agent <- left_join(cancer_agent, select(patient, PERSON_ID, COHORT_START_DATE), by = 'PERSON_ID')
cancer_agent <- unique(cancer_agent) 

afterICI_cancer_agent <- cancer_agent %>%
  mutate(
    COHORT_START_DATE = as.Date(COHORT_START_DATE),
    DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE)
  ) %>%
  filter(DRUG_EXPOSURE_START_DATE >= COHORT_START_DATE)

afterICI_cancer_agent <- select(afterICI_cancer_agent, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE)
afterICI_cancer_agent <- unique(afterICI_cancer_agent)

afterICI_cancer_agent <- afterICI_cancer_agent %>%
  group_by(PERSON_ID) %>%
  summarise(other_CA_START_DATE = min(DRUG_EXPOSURE_START_DATE, na.rm = TRUE)) %>%
  ungroup() 

patient <- left_join(patient, afterICI_cancer_agent, by = 'PERSON_ID')



write.csv(afterICI_cancer_agent, file = "afterICI_cancer_agent.CSV")



# Exclude participants who are not on ICI monotherapy at the steroid index date in the TARGET cohort
patient_mono <- patient %>%
  filter(is.na(other_CA_START_DATE) | is.na(Steroid_index_date) | 
         (other_CA_START_DATE > Steroid_index_date))

patient_mono <- patient_mono %>%
  mutate(
    other_CA = ifelse(is.na(other_CA_START_DATE), 0, 1)
  )


patient_mono <- left_join(patient_mono, ICI_drug, by = 'PERSON_ID')




```




3. Covariates: Data prep for TDPS Matching
# Time-fixed: before Cohort Entry (ICI New use)
# Time-dependent covariates:
## Cases: evaluate prescriptions using Index_ICI_days (maximum of 1 to 7 windows)
## Controls: evaluate prescriptions using ICI_use_days (same window definition)
## For each prescription window count, iterate through the loop below


```{r}

#1) FIXED data prep------------------------------------------------------------------------------------------------

#1-1) SEX / AGE
# 8532: Female
# 8507: Male
sexbirth_c <- left_join(sexbirth, select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by = "PERSON_ID")
sexbirth_c <- sexbirth_c %>%
  mutate(
    birth_date = make_date(YEAR_OF_BIRTH, MONTH_OF_BIRTH),
    COHORT_START_DATE = as.Date(COHORT_START_DATE),
    age = year(COHORT_START_DATE) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0)
  )

sexbirth_c <- sexbirth_c %>%
  mutate(
    sex = case_when(
      GENDER_CONCEPT_ID == 8507 ~ "M",
      GENDER_CONCEPT_ID == 8532 ~ "F",
      TRUE ~ NA_character_  # Assign NA when GENDER_CONCEPT_ID is not mapped to male or female
    )
  )
sexbirth_c <- sexbirth_c %>%
  filter(!is.na(TARGET))

patient_mono <- left_join(patient_mono, select(sexbirth_c, PERSON_ID, age, sex), by = 'PERSON_ID')



patient_mono <- left_join(patient_mono, autoimmunedz, by = 'PERSON_ID')

patient_mono <- left_join(patient_mono, comorbd_disease, by = 'PERSON_ID')



concom_drug <- left_join(select(patient_mono, PERSON_ID), concom_drug2, by = 'PERSON_ID')
concom_drug <- concom_drug %>%
  mutate(across(everything(), ~replace_na(., 0)))
                          
patient_mono <- left_join(patient_mono, concom_drug, by = 'PERSON_ID')                      



cancer_lung_sq <- read.csv("cancer_lung_sq.CSV")[,-1] 
patient_mono <- patient_mono %>%
  mutate(squamouslc = ifelse(PERSON_ID %in% cancer_lung_sq$PERSON_ID, 1, 0))





# baseline/sensitivity: pre-steroid 

pre_steroid <- left_join(select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), select(steroid, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE), by ='PERSON_ID') 
pre_steroid <- unique(pre_steroid)
pre_steroid <- pre_steroid %>%
  filter(!is.na(TARGET))

pre_steroid <- pre_steroid %>%
  mutate(
    DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  ) %>%
  group_by(PERSON_ID) %>%
  mutate(
    pre_steroid = if_else(any(DRUG_EXPOSURE_START_DATE <= COHORT_START_DATE & 
                              DRUG_EXPOSURE_START_DATE >= COHORT_START_DATE - 180), 1, 0)
  ) %>%
  ungroup()

pre_steroid <- select(pre_steroid, PERSON_ID, pre_steroid)
pre_steroid <- unique(pre_steroid)
pre_steroid <- pre_steroid %>%
  mutate(across(everything(), ~replace_na(., 0)))

patient_mono <- left_join(patient_mono, pre_steroid, by = 'PERSON_ID')                      





# baseline: pre-chemo

pre_chemo <- left_join(select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), select(chemo_agent, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE), by ='PERSON_ID') 
pre_chemo <- unique(pre_chemo)
pre_chemo <- pre_chemo %>%
  filter(!is.na(TARGET))

pre_chemo <- pre_chemo %>%
  mutate(
    DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  ) %>%
  group_by(PERSON_ID) %>%
  mutate(
    pre_chemo = if_else(any(DRUG_EXPOSURE_START_DATE <= COHORT_START_DATE & 
                              DRUG_EXPOSURE_START_DATE >= COHORT_START_DATE - 180), 1, 0)
  ) %>%
  ungroup()

pre_chemo <- select(pre_chemo, PERSON_ID, pre_chemo)
pre_chemo <- unique(pre_chemo)
pre_chemo <- pre_chemo %>%
  mutate(across(everything(), ~replace_na(., 0)))

patient_mono <- left_join(patient_mono, pre_chemo, by = 'PERSON_ID')                      




pre_TKI_ALK <- left_join(select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), select(TKI_ALK, PERSON_ID, DRUG_EXPOSURE_START_DATE, DRUG_EXPOSURE_END_DATE), by ='PERSON_ID') 
pre_TKI_ALK <- unique(pre_TKI_ALK)
pre_TKI_ALK <- pre_TKI_ALK %>%
  filter(!is.na(TARGET))

pre_TKI_ALK <- pre_TKI_ALK %>%
  mutate(
    DRUG_EXPOSURE_START_DATE = as.Date(DRUG_EXPOSURE_START_DATE),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  ) %>%
  group_by(PERSON_ID) %>%
  mutate(
    pre_TKI_ALK = if_else(any(DRUG_EXPOSURE_START_DATE <= COHORT_START_DATE & 
                              DRUG_EXPOSURE_START_DATE >= COHORT_START_DATE - 180), 1, 0)
  ) %>%
  ungroup()

pre_TKI_ALK <- select(pre_TKI_ALK, PERSON_ID, pre_TKI_ALK)
pre_TKI_ALK <- unique(pre_TKI_ALK)
pre_TKI_ALK <- pre_TKI_ALK %>%
  mutate(across(everything(), ~replace_na(., 0)))

patient_mono <- left_join(patient_mono, pre_TKI_ALK, by = 'PERSON_ID')                      





patient_mono <- patient_mono %>%
  rowwise() %>%
  mutate(
    ICI_dupl = if_else(sum(pembrolizumab_use, nivolumab_use, atezolizumab_use, durvalumab_use) >= 2, 1, 0)
  ) %>%
  ungroup()













#2) Time-dependent data prep-------------------------------------------------------------------------

#2-1)

patient_mono <- patient_mono %>%
  mutate(
    ad_steroid = case_when(
      TARGET == 1 & Index_ICI_days < 30 ~ 1,
      TARGET == 1 & Index_ICI_days >= 30 & Index_ICI_days < 60 ~ 2,
      TARGET == 1 & Index_ICI_days >= 60 & Index_ICI_days < 120 ~ 3,
      TARGET == 1 & Index_ICI_days >= 120 & Index_ICI_days < 180 ~ 4,
      TARGET == 1 & Index_ICI_days >= 180 & Index_ICI_days < 270 ~ 5,
      TARGET == 1 & Index_ICI_days >= 270 & Index_ICI_days < 365 ~ 6,
      TARGET == 1 & Index_ICI_days >= 365 & Index_ICI_days < 545 ~ 7,
      TARGET == 1 & Index_ICI_days >= 545 & Index_ICI_days < 790 ~ 8,
      TARGET == 1 & Index_ICI_days >= 790 ~ 9))




patient_mono <- patient_mono %>%
  mutate(
    ICI_prescrib = case_when(
      TARGET == 1 & ICI_use_days < 30 ~ 1,
      TARGET == 1 & ICI_use_days >= 30 & ICI_use_days < 60 ~ 2,
      TARGET == 1 & ICI_use_days >= 60 & ICI_use_days < 120 ~ 3,
      TARGET == 1 & ICI_use_days >= 120 & ICI_use_days < 180 ~ 4,
      TARGET == 1 & ICI_use_days >= 180 & ICI_use_days < 270 ~ 5,
      TARGET == 1 & ICI_use_days >= 270 & ICI_use_days < 365 ~ 6,
      TARGET == 1 & ICI_use_days >= 365 & ICI_use_days < 545 ~ 7,
      TARGET == 1 & ICI_use_days >= 545 & ICI_use_days < 730 ~ 8,
      TARGET == 1 & ICI_use_days >= 730 ~ 9,
      TARGET == 0 & ICI_use_days < 30 ~ 1,
      TARGET == 0 & ICI_use_days >= 30 & ICI_use_days < 60 ~ 2,
      TARGET == 0 & ICI_use_days >= 60 & ICI_use_days < 120 ~ 3,
      TARGET == 0 & ICI_use_days >= 120 & ICI_use_days < 180 ~ 4,
      TARGET == 0 & ICI_use_days >= 180 & ICI_use_days < 270 ~ 5,
      TARGET == 0 & ICI_use_days >= 270 & ICI_use_days < 365 ~ 6,
      TARGET == 0 & ICI_use_days >= 365 & ICI_use_days < 545 ~ 7,
      TARGET == 0 & ICI_use_days >= 545 & ICI_use_days < 730 ~ 8,
      TARGET == 0 & ICI_use_days >= 730 ~ 9
    )
  )


patient_mono <- patient_mono %>%
  mutate(
    ICI_prescrib_TDPS = pmax(ICI_prescrib, ad_steroid, na.rm = TRUE)
  )




patient_mono_time <- patient_mono %>%
  uncount(weights = ICI_prescrib_TDPS, .remove = FALSE)  


patient_mono_time <- patient_mono_time %>%
  group_by(PERSON_ID) %>%  
  mutate(ICI_prescrib_TDPS = row_number()) %>% 
  ungroup() 




#2-2)


###

#Sepsis

sepsis <- left_join(sepsis, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

sepsis <- sepsis %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_sepsis_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()


sepsis <- sepsis %>%
  mutate(
    Sepsis_occur = case_when(
      Index_sepsis_days >= -365 & Index_sepsis_days < 0 ~ 1,
      Index_sepsis_days >= 0 & Index_sepsis_days < 30 ~ 2,
      Index_sepsis_days >= 30 & Index_sepsis_days < 60 ~ 3,
      Index_sepsis_days >= 60 & Index_sepsis_days < 120 ~ 4,
      Index_sepsis_days >= 120 & Index_sepsis_days < 180 ~ 5,
      Index_sepsis_days >= 180 & Index_sepsis_days < 270 ~ 6,
      Index_sepsis_days >= 270 & Index_sepsis_days < 365~ 7,
      Index_sepsis_days >= 365 & Index_sepsis_days < 545~ 8,
      Index_sepsis_days >= 545 & Index_sepsis_days < 730~ 9
    )
  )



sepsis <- sepsis %>%
  group_by(PERSON_ID, Sepsis_occur) %>%
  filter(CONDITION_START_DATE == max(CONDITION_START_DATE)) %>%
  ungroup()

patient_mono_time <- left_join(patient_mono_time, select(sepsis, c('PERSON_ID', 'Sepsis_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(Sepsis = if_else(is.na(Sepsis_occur), 
                          0, 
                          if_else(Sepsis_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -Sepsis_occur)
patient_mono_time <- unique(patient_mono_time)




#Pneumonitis
pneumonitis <- left_join(pneumonitis, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

pneumonitis <- pneumonitis %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_pneumonitis_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

pneumonitis <- pneumonitis %>%
  mutate(
    Pneumonitis_occur = case_when(
      Index_pneumonitis_days >= -365 & Index_pneumonitis_days < 0 ~ 1,
      Index_pneumonitis_days >= 0 & Index_pneumonitis_days < 30 ~ 2,
      Index_pneumonitis_days >= 30 & Index_pneumonitis_days < 60 ~ 3,
      Index_pneumonitis_days >= 60 & Index_pneumonitis_days < 120 ~ 4,
      Index_pneumonitis_days >= 120 & Index_pneumonitis_days < 180 ~ 5,
      Index_pneumonitis_days >= 180 & Index_pneumonitis_days < 270 ~ 6,
      Index_pneumonitis_days >= 270 & Index_pneumonitis_days < 365 ~ 7,
      Index_pneumonitis_days >= 365 & Index_pneumonitis_days < 545 ~ 8,
      Index_pneumonitis_days >= 545 & Index_pneumonitis_days < 730 ~ 9
    )
  )

pneumonitis <- pneumonitis %>%
  group_by(PERSON_ID, Pneumonitis_occur) %>%
  filter(CONDITION_START_DATE == max(CONDITION_START_DATE)) %>%
  ungroup()

pneumonitis <- pneumonitis %>%
  group_by(PERSON_ID) %>%
  filter(Pneumonitis_occur == min(Pneumonitis_occur)) %>%
  ungroup()

pneumonitis <- pneumonitis %>%
  distinct(PERSON_ID, .keep_all = TRUE)

pneumonitis <- na.omit(pneumonitis, subset = "Pneumonitis_occur")

patient_mono_time <- left_join(patient_mono_time, select(pneumonitis, c('PERSON_ID', 'Pneumonitis_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(Pneumonitis = if_else(is.na(Pneumonitis_occur), 
                          0, 
                          if_else(Pneumonitis_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -Pneumonitis_occur)




#Dyspnea
dyspnea <- left_join(dyspnea, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

dyspnea <- dyspnea %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_dyspnea_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

dyspnea <- dyspnea %>%
  mutate(
    dyspnea_occur = case_when(
      Index_dyspnea_days >= -365 & Index_dyspnea_days < 0 ~ 1,
      Index_dyspnea_days >= 0 & Index_dyspnea_days < 30 ~ 2,
      Index_dyspnea_days >= 30 & Index_dyspnea_days < 60 ~ 3,
      Index_dyspnea_days >= 60 & Index_dyspnea_days < 120 ~ 4,
      Index_dyspnea_days >= 120 & Index_dyspnea_days < 180 ~ 5,
      Index_dyspnea_days >= 180 & Index_dyspnea_days < 270 ~ 6,
      Index_dyspnea_days >= 270 & Index_dyspnea_days < 365 ~ 7,
      Index_dyspnea_days >= 365 & Index_dyspnea_days < 545 ~ 8,
      Index_dyspnea_days >= 545 & Index_dyspnea_days < 730 ~ 9
    )
  )

dyspnea <- dyspnea %>%
  group_by(PERSON_ID, dyspnea_occur) %>%
  filter(CONDITION_START_DATE == max(CONDITION_START_DATE)) %>%
  ungroup()

dyspnea$ICI_prescrib_TDPS <- dyspnea$dyspnea_occur

dyspnea <- dyspnea %>%
  distinct(PERSON_ID, dyspnea_occur, .keep_all = TRUE)

dyspnea <- na.omit(dyspnea, subset = "dyspnea_occur")


patient_mono_time <- left_join(patient_mono_time, select(dyspnea, c('PERSON_ID', 'dyspnea_occur','ICI_prescrib_TDPS')), by = c('PERSON_ID','ICI_prescrib_TDPS'))
patient_mono_time <- patient_mono_time %>%
  mutate(Dyspnea = if_else(is.na(dyspnea_occur), 
                          0, 
                          1))
patient_mono_time <- select(patient_mono_time, -dyspnea_occur)





# RT_GKS
CSC_RT_GKS <- left_join(CSC_RT_GKS, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')
RT_GKS <- CSC_RT_GKS %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_RT_days = as.numeric(difftime(PROCEDURE_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

RT_GKS <- RT_GKS %>%
  mutate(
    RT_occur = case_when(
      Index_RT_days >= -180 & Index_RT_days < 0 ~ 1,
      Index_RT_days >= 0 & Index_RT_days < 30 ~ 2,
      Index_RT_days >= 30 & Index_RT_days < 60 ~ 3,
      Index_RT_days >= 60 & Index_RT_days < 120 ~ 4,
      Index_RT_days >= 120 & Index_RT_days < 180 ~ 5,
      Index_RT_days >= 180 & Index_RT_days < 270 ~ 6,
      Index_RT_days >= 270 & Index_RT_days < 365 ~ 7,
      Index_RT_days >= 365 & Index_RT_days < 545 ~ 8,
      Index_RT_days >= 545 & Index_RT_days < 730 ~ 9
    )
  )

RT_GKS <- RT_GKS %>%
  group_by(PERSON_ID, RT_occur) %>%
  filter(PROCEDURE_DATE == max(PROCEDURE_DATE)) %>%
  ungroup()

RT_GKS$ICI_prescrib_TDPS <- RT_GKS$RT_occur

RT_GKS <- RT_GKS %>%
  distinct(PERSON_ID, RT_occur, .keep_all = TRUE)

RT_GKS <- na.omit(RT_GKS, subset = "RT_occur")


patient_mono_time <- left_join(patient_mono_time, select(RT_GKS, c('PERSON_ID', 'RT_occur','ICI_prescrib_TDPS')), by = c('PERSON_ID','ICI_prescrib_TDPS'))
patient_mono_time <- patient_mono_time %>%
  mutate(RT_GKS = if_else(is.na(RT_occur), 
                          0, 
                          1))
patient_mono_time <- select(patient_mono_time, -RT_occur)






# Brain metastasis
brainmeta <- left_join(brainmeta, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

brainmeta <- brainmeta %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_brainmeta_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

brainmeta <- brainmeta %>%
  mutate(
    brainmeta_occur = case_when(
      Index_brainmeta_days >= -365 & Index_brainmeta_days < 0 ~ 1,
      Index_brainmeta_days >= 0 & Index_brainmeta_days < 30 ~ 2,
      Index_brainmeta_days >= 30 & Index_brainmeta_days < 60 ~ 3,
      Index_brainmeta_days >= 60 & Index_brainmeta_days < 120 ~ 4,
      Index_brainmeta_days >= 120 & Index_brainmeta_days < 180 ~ 5,
      Index_brainmeta_days >= 180 & Index_brainmeta_days < 270 ~ 6,
      Index_brainmeta_days >= 270 & Index_brainmeta_days < 365 ~ 7,
      Index_brainmeta_days >= 365 & Index_brainmeta_days < 545 ~ 8,
      Index_brainmeta_days >= 545 & Index_brainmeta_days < 730 ~ 9
    )
  )

brainmeta <- brainmeta %>%
  group_by(PERSON_ID, brainmeta_occur) %>%
  filter(CONDITION_START_DATE == min(CONDITION_START_DATE)) %>%
  ungroup()

brainmeta <- brainmeta %>%
  group_by(PERSON_ID) %>%
  filter(brainmeta_occur == min(brainmeta_occur)) %>%
  ungroup()

brainmeta <- brainmeta %>%
  distinct(PERSON_ID, .keep_all = TRUE)

brainmeta <- na.omit(brainmeta, subset = "brainmeta_occur")

patient_mono_time <- left_join(patient_mono_time, select(brainmeta, c('PERSON_ID', 'brainmeta_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(Brainmeta = if_else(is.na(brainmeta_occur), 
                          0, 
                          if_else(brainmeta_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -brainmeta_occur)






#Bone metastasis
bonemeta <- left_join(bonemeta, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

bonemeta <- bonemeta %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_bonemeta_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

bonemeta <- bonemeta %>%
  mutate(
    bonemeta_occur = case_when(
      Index_bonemeta_days >= -365 & Index_bonemeta_days < 0 ~ 1,
      Index_bonemeta_days >= 0 & Index_bonemeta_days < 30 ~ 2,
      Index_bonemeta_days >= 30 & Index_bonemeta_days < 60 ~ 3,
      Index_bonemeta_days >= 60 & Index_bonemeta_days < 120 ~ 4,
      Index_bonemeta_days >= 120 & Index_bonemeta_days < 180 ~ 5,
      Index_bonemeta_days >= 180 & Index_bonemeta_days < 270 ~ 6,
      Index_bonemeta_days >= 270 & Index_bonemeta_days < 365 ~ 7,
      Index_bonemeta_days >= 365 & Index_bonemeta_days < 545 ~ 8,
      Index_bonemeta_days >= 545 & Index_bonemeta_days < 730 ~ 9
    )
  )

bonemeta <- bonemeta %>%
  group_by(PERSON_ID, bonemeta_occur) %>%
  filter(CONDITION_START_DATE == min(CONDITION_START_DATE)) %>%
  ungroup()

bonemeta <- bonemeta %>%
  group_by(PERSON_ID) %>%
  filter(bonemeta_occur == min(bonemeta_occur)) %>%
  ungroup()

bonemeta <- bonemeta %>%
  distinct(PERSON_ID, .keep_all = TRUE)

bonemeta <- na.omit(bonemeta, subset = "bonemeta_occur")

patient_mono_time <- left_join(patient_mono_time, select(bonemeta, c('PERSON_ID', 'bonemeta_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(Bonemeta = if_else(is.na(bonemeta_occur), 
                          0, 
                          if_else(bonemeta_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -bonemeta_occur)





#Liver metastasis
livermeta <- left_join(livermeta, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

livermeta <- livermeta %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_livermeta_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

livermeta <- livermeta %>%
  mutate(
    livermeta_occur = case_when(
      Index_livermeta_days >= -365 & Index_livermeta_days < 0 ~ 1,
      Index_livermeta_days >= 0 & Index_livermeta_days < 30 ~ 2,
      Index_livermeta_days >= 30 & Index_livermeta_days < 60 ~ 3,
      Index_livermeta_days >= 60 & Index_livermeta_days < 120 ~ 4,
      Index_livermeta_days >= 120 & Index_livermeta_days < 180 ~ 5,
      Index_livermeta_days >= 180 & Index_livermeta_days < 270 ~ 6,
      Index_livermeta_days >= 270 & Index_livermeta_days < 365 ~ 7,
      Index_livermeta_days >= 365 & Index_livermeta_days < 545 ~ 8,
      Index_livermeta_days >= 545 & Index_livermeta_days < 730 ~ 9
    )
  )

livermeta <- livermeta %>%
  group_by(PERSON_ID, livermeta_occur) %>%
  filter(CONDITION_START_DATE == min(CONDITION_START_DATE)) %>%
  ungroup()

livermeta <- livermeta %>%
  group_by(PERSON_ID) %>%
  filter(livermeta_occur == min(livermeta_occur)) %>%
  ungroup()

livermeta <- livermeta %>%
  distinct(PERSON_ID, .keep_all = TRUE)

livermeta <- na.omit(livermeta, subset = "livermeta_occur")

patient_mono_time <- left_join(patient_mono_time, select(livermeta, c('PERSON_ID', 'livermeta_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(Livermeta = if_else(is.na(livermeta_occur), 
                          0, 
                          if_else(livermeta_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -livermeta_occur)






# Cancer registry and metastasis_all integratio
cancer_reg <- cancer_reg_final[,-(2:4)]
colnames(cancer_reg)
cancer_reg <- unique(cancer_reg)
cancer_reg <- cancer_reg %>%
  filter(is.na(VALUE_AS_STRING.y) != TRUE | is.na(VALUE_AS_STRING) != TRUE)
cancer_reg <- cancer_reg[!grepl("M0", cancer_reg$VALUE_AS_STRING.y), ]
cancer_reg <- cancer_reg %>%
  mutate(reg_meta_history = 1)
cancer_reg <- cancer_reg[,-(2:4)]
cancer_reg <- unique(cancer_reg)

cancer_reg <- cancer_reg %>%
  group_by(PERSON_ID) %>%
  mutate(
    meta1 = ifelse(OBSERVATION_SOURCE_VALUE == "metacode1", VALUE_AS_STRING, NA),
    meta2 = ifelse(OBSERVATION_SOURCE_VALUE == "metacode2", VALUE_AS_STRING, NA),
    meta3 = ifelse(OBSERVATION_SOURCE_VALUE == "metacode3", VALUE_AS_STRING, NA)
  ) %>%
  ungroup()
cancer_reg <- cancer_reg[,-(2:4)]
cancer_reg <- unique(cancer_reg)
cancer_reg <- cancer_reg %>%
  group_by(PERSON_ID) %>%
  summarise(
    meta1 = na.omit(meta1)[1],
    meta2 = na.omit(meta2)[1],
    meta3 = na.omit(meta3)[1],
    reg_meta_history = first(reg_meta_history)
  ) %>%
  ungroup()




metastasis_kcd <- read.csv("metastasis_all_kcd.csv")
metastasis_kcd <- unique(metastasis_kcd)


metastasis_kcd <- metastasis_kcd %>%
  group_by(condition_concept_id) %>%
  filter(
    if(any(ext_cond_source_value_kcd != "NULL")){
      ext_cond_source_value_kcd != "NULL"
    }else{
      TRUE
    }
  ) %>% ungroup()

metastasis_kcd <- metastasis_kcd %>%
  mutate(meta_source = substr(ext_cond_source_value_kcd, 1, 4))
metastasis_kcd <- metastasis_kcd[,-(2:3)]
metastasis_kcd <- unique(metastasis_kcd)


## Metastases corresponding to M1a~b
# 254591	94391008	Secondary malignant neoplasm of lung
# 72266	94493005	Secondary malignant neoplasm of pleura
# 4248199	94484009	Secondary malignant neoplasm of pericardium
# 439751	94351005	Secondary malignant neoplasm of intrathoracic lymph nodes
# 4248061	94352003	Secondary malignant neoplasm of intrathoracic organs
# 4312284	94233006	Secondary malignant neoplasm of bronchus




colnames(cancer_primary)
cancer_primary <- unique(select(cancer_primary, PERSON_ID, CONDITION_CONCEPT_ID, CONDITION_START_DATE, CONDITION_END_DATE, EXT_COND_SOURCE_VALUE_KCD))
cancer_primary <- cancer_primary %>%
  group_by(PERSON_ID, CONDITION_CONCEPT_ID) %>%
  slice_min(CONDITION_START_DATE) %>%
  ungroup()
#### Confirmed that all non-lung primary cancers were excluded




##Use the "metastasis_all" dataset for the new metastasis outcome; compare the earliest date per concept_id with the index date.



metastasis_all  <- left_join(metastasis_all , select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')



metastasis <- metastasis_all %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_metastasis_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

metastasis <- metastasis %>%
  mutate(
    metastasis_occur = case_when(
      Index_metastasis_days >= -365 & Index_metastasis_days < 0 ~ 1,
      Index_metastasis_days >= 0 & Index_metastasis_days < 30 ~ 2,
      Index_metastasis_days >= 30 & Index_metastasis_days < 60 ~ 3,
      Index_metastasis_days >= 60 & Index_metastasis_days < 120 ~ 4,
      Index_metastasis_days >= 120 & Index_metastasis_days < 180 ~ 5,
      Index_metastasis_days >= 180 & Index_metastasis_days < 270 ~ 6,
      Index_metastasis_days >= 270 & Index_metastasis_days < 365 ~ 7,
      Index_metastasis_days >= 365 & Index_metastasis_days < 545 ~ 8,
      Index_metastasis_days >= 545 & Index_metastasis_days < 730 ~ 9
    )
  )

metastasis <- metastasis %>%
  group_by(PERSON_ID, metastasis_occur) %>%
  filter(CONDITION_START_DATE == min(CONDITION_START_DATE)) %>%
  ungroup()

metastasis <- metastasis %>%
  group_by(PERSON_ID) %>%
  filter(metastasis_occur == min(metastasis_occur)) %>%
  ungroup()

metastasis <- metastasis %>%
  distinct(PERSON_ID, .keep_all = TRUE)

metastasis <- na.omit(metastasis, subset = "metastasis_occur")


metastasis <- left_join(metastasis, select(cancer_reg, PERSON_ID, reg_meta_history), by = "PERSON_ID")
metastasis <- metastasis %>%
  mutate(metastasis_occur_new = metastasis_occur)
metastasis <- metastasis %>%
  mutate(metastasis_occur = ifelse(!is.na(reg_meta_history), 1, metastasis_occur))


patient_mono_time <- left_join(patient_mono_time, select(metastasis, c('PERSON_ID', 'metastasis_occur')), by = 'PERSON_ID')
patient_mono_time <- patient_mono_time %>%
  mutate(metastasis = if_else(is.na(metastasis_occur), 
                          0, 
                          if_else(metastasis_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -metastasis_occur)








# Age
patient_mono_time <- left_join(patient_mono_time, select(sexbirth, c(PERSON_ID, YEAR_OF_BIRTH, MONTH_OF_BIRTH)), by = 'PERSON_ID')

patient_mono_time <- patient_mono_time %>%
  mutate(
    age = case_when(
      ICI_prescrib_TDPS == 1 ~ year(COHORT_START_DATE) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 2 ~ year(COHORT_START_DATE+days(30)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 3 ~ year(COHORT_START_DATE+days(60)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 4 ~ year(COHORT_START_DATE+days(120)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 5 ~ year(COHORT_START_DATE+days(180)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 6 ~ year(COHORT_START_DATE+days(270)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 7 ~ year(COHORT_START_DATE+days(365)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 8 ~ year(COHORT_START_DATE+days(545)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0),
      ICI_prescrib_TDPS == 9 ~ year(COHORT_START_DATE+days(730)) - YEAR_OF_BIRTH - if_else(month(COHORT_START_DATE) < MONTH_OF_BIRTH | (month(COHORT_START_DATE) == MONTH_OF_BIRTH & day(COHORT_START_DATE) < 1), 1, 0)
    )
  )



patient_mono_time <- select(patient_mono_time, -c(YEAR_OF_BIRTH, MONTH_OF_BIRTH))









#1-2) BMI
height <- left_join(height, select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
height <- unique(height)
height <- height %>%
  filter(!is.na(TARGET))
height <- height %>%
  mutate(
    MEASUREMENT_DATE = as.Date(MEASUREMENT_DATE),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  )

weight <- left_join(weight, select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
weight <- unique(weight)
weight <- weight %>%
  filter(!is.na(TARGET))
weight <- weight %>%
  mutate(
    MEASUREMENT_DATE = as.Date(MEASUREMENT_DATE),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  )


BMI <- left_join(select(patient_mono, PERSON_ID), select(weight, PERSON_ID, MEASUREMENT_DATE, VALUE_AS_NUMBER), by = 'PERSON_ID')
BMI <- full_join(BMI, select(height, PERSON_ID, MEASUREMENT_DATE, VALUE_AS_NUMBER), by = 'PERSON_ID')
BMI$MEASUREMENT_DATE.x <- as.Date(BMI$MEASUREMENT_DATE.x)
BMI$MEASUREMENT_DATE.y <- as.Date(BMI$MEASUREMENT_DATE.y)

# Filter rows where the dates are exactly the same
BMI <- BMI[abs(difftime(BMI$MEASUREMENT_DATE.x, BMI$MEASUREMENT_DATE.y, units = "days")) < 90, ]
BMI <- unique(BMI)

BMI <- BMI %>%
  mutate(BMI_cal = ifelse(!is.na(VALUE_AS_NUMBER.y) | !is.na(VALUE_AS_NUMBER.x), VALUE_AS_NUMBER.x / ((VALUE_AS_NUMBER.y/100)^2), NA))



BMI <- left_join(BMI, select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
BMI <- BMI %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_BMI_days = as.numeric(difftime(MEASUREMENT_DATE.x, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

BMI <- BMI %>%
  mutate(
    BMI_occur = case_when(
      Index_BMI_days >= -365 & Index_BMI_days < 0 ~ 1,
      Index_BMI_days >= 0 & Index_BMI_days < 30 ~ 2,
      Index_BMI_days >= 30 & Index_BMI_days < 60 ~ 3,
      Index_BMI_days >= 60 & Index_BMI_days < 120 ~ 4,
      Index_BMI_days >= 120 & Index_BMI_days < 180 ~ 5,
      Index_BMI_days >= 180 & Index_BMI_days < 270 ~ 6,
      Index_BMI_days >= 270 & Index_BMI_days < 365 ~ 7,
      Index_BMI_days >= 365 & Index_BMI_days < 545 ~ 8,
      Index_BMI_days >= 545 & Index_BMI_days < 730 ~ 9
    )
  )

BMI <- BMI %>%
  group_by(PERSON_ID, BMI_occur) %>%
  filter(MEASUREMENT_DATE.x == max(MEASUREMENT_DATE.x)) %>%
  ungroup()

BMI <- BMI %>%
  group_by(PERSON_ID, BMI_occur) %>%
  filter(abs(difftime(MEASUREMENT_DATE.x, MEASUREMENT_DATE.y, units = "days")) == 
           min(abs(difftime(MEASUREMENT_DATE.x, MEASUREMENT_DATE.y, units = "days")), na.rm = TRUE)) %>%
  ungroup()

BMI <- unique(BMI)

BMI <- BMI %>%
  filter(!is.na(BMI_occur))

BMI <- BMI %>%
  distinct(PERSON_ID, BMI_occur, .keep_all = TRUE)


BMI$ICI_prescrib_TDPS <- BMI$BMI_occur
###250318 weight 변수 추가
patient_mono_time <- left_join(patient_mono_time, select(BMI, PERSON_ID, ICI_prescrib_TDPS, BMI_cal, VALUE_AS_NUMBER.x), by = c('PERSON_ID', 'ICI_prescrib_TDPS'))

patient_mono_time <- patient_mono_time %>%
  mutate(weight = VALUE_AS_NUMBER.x)
patient_mono_time <- select(patient_mono_time, -VALUE_AS_NUMBER.x)


# 1-4) tumor marker -> Time-dependent variables
# CEA: 36313717,4244721
# cytokeratin: 3000659
# unit value: ng/mL

cytokeratin <- left_join(select(cytokeratin, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
cytokeratin <- unique(cytokeratin)
cytokeratin <- cytokeratin %>%
  filter(!is.na(TARGET))

cytokeratin <- cytokeratin %>%
  mutate(
    MEASUREMENT_DATETIME = as.Date(MEASUREMENT_DATETIME),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  )

cytokeratin <- cytokeratin %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_cytokeratin_days = as.numeric(difftime(MEASUREMENT_DATETIME, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

cytokeratin <- cytokeratin %>%
  mutate(
    cytokeratin_occur = case_when(
      Index_cytokeratin_days >= -365 & Index_cytokeratin_days < 0 ~ 1,
      Index_cytokeratin_days >= 0 & Index_cytokeratin_days < 30 ~ 2,
      Index_cytokeratin_days >= 30 & Index_cytokeratin_days < 60 ~ 3,
      Index_cytokeratin_days >= 60 & Index_cytokeratin_days < 120 ~ 4,
      Index_cytokeratin_days >= 120 & Index_cytokeratin_days < 180 ~ 5,
      Index_cytokeratin_days >= 180 & Index_cytokeratin_days < 270 ~ 6,
      Index_cytokeratin_days >= 270 & Index_cytokeratin_days < 365 ~ 7,
      Index_cytokeratin_days >= 365 & Index_cytokeratin_days < 545 ~ 8,
      Index_cytokeratin_days >= 545 & Index_cytokeratin_days < 730 ~ 9
    )
  )

cytokeratin <- cytokeratin %>%
  group_by(PERSON_ID, cytokeratin_occur) %>%
  filter(MEASUREMENT_DATETIME == max(MEASUREMENT_DATETIME)) %>%
  ungroup()

cytokeratin <- cytokeratin %>%
  filter(!is.na(cytokeratin_occur))

cytokeratin <- cytokeratin %>%
  distinct(PERSON_ID, cytokeratin_occur, .keep_all = TRUE)

cytokeratin$ICI_prescrib_TDPS <- cytokeratin$cytokeratin_occur
cytokeratin$Cytokeratin <- cytokeratin$VALUE_AS_NUMBER
patient_mono_time <- left_join(patient_mono_time, select(cytokeratin, PERSON_ID, ICI_prescrib_TDPS, Cytokeratin), by = c('PERSON_ID', 'ICI_prescrib_TDPS'))












#1-3) CBC
print(unique(CBC_lab$MEASUREMENT_CONCEPT_ID))

ANC <- CBC_lab %>%
  filter(MEASUREMENT_CONCEPT_ID == 3013650)
# unit value: /㎕

ANC <- left_join(select(ANC, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
ANC <- unique(ANC)
ANC <- ANC %>%
  filter(!is.na(TARGET))

ANC <- ANC %>%
  mutate(
    MEASUREMENT_DATETIME = as.Date(MEASUREMENT_DATETIME),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  )

ANC <- ANC %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_ANC_days = as.numeric(difftime(MEASUREMENT_DATETIME, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

ANC <- ANC %>%
  mutate(
    ANC_occur = case_when(
      Index_ANC_days >= -365 & Index_ANC_days < 0 ~ 1,
      Index_ANC_days >= 0 & Index_ANC_days < 30 ~ 2,
      Index_ANC_days >= 30 & Index_ANC_days < 60 ~ 3,
      Index_ANC_days >= 60 & Index_ANC_days < 120 ~ 4,
      Index_ANC_days >= 120 & Index_ANC_days < 180 ~ 5,
      Index_ANC_days >= 180 & Index_ANC_days < 270 ~ 6,
      Index_ANC_days >= 270 & Index_ANC_days < 365 ~ 7,
      Index_ANC_days >= 365 & Index_ANC_days < 545 ~ 8,
      Index_ANC_days >= 545 & Index_ANC_days < 730 ~ 9
    )
  )

ANC <- ANC %>%
  group_by(PERSON_ID, ANC_occur) %>%
  filter(MEASUREMENT_DATETIME == max(MEASUREMENT_DATETIME)) %>%
  ungroup()

ANC <- ANC %>%
  filter(!is.na(ANC_occur))

ANC <- ANC %>%
  distinct(PERSON_ID, ANC_occur, .keep_all = TRUE)

ANC$ICI_prescrib_TDPS <- ANC$ANC_occur
ANC$ANC_lab <- ANC$VALUE_AS_NUMBER
patient_mono_time <- left_join(patient_mono_time, select(ANC, PERSON_ID, ICI_prescrib_TDPS, ANC_lab), by = c('PERSON_ID', 'ICI_prescrib_TDPS'))











#NLR

Lymphocyte <- CBC_lab %>%
  filter(MEASUREMENT_CONCEPT_ID == 3037511)
# unit value: %
Neutrophil <- CBC_lab %>%
  filter(MEASUREMENT_CONCEPT_ID == 3017354)
# unit value: %


Lymphocyte <- left_join(select(Lymphocyte, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
Lymphocyte <- unique(Lymphocyte)
Lymphocyte <- Lymphocyte %>%
  filter(!is.na(TARGET))
Lymphocyte <- Lymphocyte %>%
  mutate(
    MEASUREMENT_DATETIME = as.Date(MEASUREMENT_DATETIME),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  ) 

Neutrophil <- left_join(select(Neutrophil, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
Neutrophil <- unique(Neutrophil)
Neutrophil <- Neutrophil %>%
  filter(!is.na(TARGET))
Neutrophil <- Neutrophil %>%
  mutate(
    MEASUREMENT_DATETIME = as.Date(MEASUREMENT_DATETIME),
    COHORT_START_DATE = as.Date(COHORT_START_DATE)
  )


NLR <- left_join(select(patient_mono, PERSON_ID), select(Lymphocyte, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), by = 'PERSON_ID')
NLR <- full_join(NLR, select(Neutrophil, PERSON_ID, MEASUREMENT_DATETIME, VALUE_AS_NUMBER), by = 'PERSON_ID')
NLR$MEASUREMENT_DATETIME.x <- as.Date(NLR$MEASUREMENT_DATETIME.x)
NLR$MEASUREMENT_DATETIME.y <- as.Date(NLR$MEASUREMENT_DATETIME.y)

# Filter rows where the dates are exactly the same
# NLR <- NLR[abs(difftime(NLR$MEASUREMENT_DATETIME.x, NLR$MEASUREMENT_DATETIME.y, units = "days")) < 3, ]
NLR <- NLR[NLR$MEASUREMENT_DATETIME.x == NLR$MEASUREMENT_DATETIME.y, ]

NLR <- unique(NLR)

NLR <- NLR %>%
  mutate(NLR_cal = ifelse(!is.na(VALUE_AS_NUMBER.y) | !is.na(VALUE_AS_NUMBER.x), (VALUE_AS_NUMBER.y/VALUE_AS_NUMBER.x), NA))
#Neutrophil/Lymphocyt


NLR <- left_join(NLR, select(patient_mono, PERSON_ID, COHORT_START_DATE, TARGET), by ='PERSON_ID')
NLR <- NLR %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_NLR_days = as.numeric(difftime(MEASUREMENT_DATETIME.x, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

NLR <- NLR %>%
  mutate(
    NLR_occur = case_when(
      Index_NLR_days >= -365 & Index_NLR_days < 0 ~ 1,
      Index_NLR_days >= 0 & Index_NLR_days < 30 ~ 2,
      Index_NLR_days >= 30 & Index_NLR_days < 60 ~ 3,
      Index_NLR_days >= 60 & Index_NLR_days < 120 ~ 4,
      Index_NLR_days >= 120 & Index_NLR_days < 180 ~ 5,
      Index_NLR_days >= 180 & Index_NLR_days < 270 ~ 6,
      Index_NLR_days >= 270 & Index_NLR_days < 365 ~ 7,
      Index_NLR_days >= 365 & Index_NLR_days < 545 ~ 8,
      Index_NLR_days >= 545 & Index_NLR_days < 730 ~ 9
    )
  )

NLR <- NLR %>%
  group_by(PERSON_ID, NLR_occur) %>%
  filter(MEASUREMENT_DATETIME.x == max(MEASUREMENT_DATETIME.x)) %>%
  ungroup()

NLR <- unique(NLR)

NLR <- NLR %>%
  filter(!is.na(NLR_occur))

NLR <- NLR %>%
  distinct(PERSON_ID, NLR_occur, .keep_all = TRUE)


NLR$ICI_prescrib_TDPS <- NLR$NLR_occur
patient_mono_time <- left_join(patient_mono_time, select(NLR, PERSON_ID, ICI_prescrib_TDPS, NLR_cal), by = c('PERSON_ID', 'ICI_prescrib_TDPS'))





# ICI period
ICI_cycle$DRUG_EXPOSURE_START_DATE <- as.Date(ICI_cycle$DRUG_EXPOSURE_START_DATE)
ICI_cycle$DRUG_EXPOSURE_END_DATE <- as.Date(ICI_cycle$DRUG_EXPOSURE_END_DATE)


ICI_period <- left_join(select(patient_mono, PERSON_ID, COHORT_START_DATE), ICI_cycle, by = "PERSON_ID")

ICI_period <- ICI_period %>%
  arrange(PERSON_ID, DRUG_EXPOSURE_START_DATE) %>%
  group_by(PERSON_ID) %>%
  mutate(
    next_start_date = lead(DRUG_EXPOSURE_START_DATE),
    days_between = if_else(
      is.na(next_start_date), 
      NA_real_, 
      as.numeric(next_start_date - DRUG_EXPOSURE_END_DATE)
    ),
    gap_over_60 = if_else(!is.na(days_between) & days_between >= 60 & days_between < 120, TRUE, FALSE)
  ) %>%
  ungroup()


ICI_period <- ICI_period %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Day_From_Start = as.numeric(difftime(DRUG_EXPOSURE_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

ICI_period <- ICI_period %>%
  mutate(
    ICI_prescrib_TDPS = case_when(
      Day_From_Start < 30 ~ 1,
      Day_From_Start >= 30 & Day_From_Start < 60 ~ 2,
      Day_From_Start >= 60 & Day_From_Start < 120 ~ 3,
      Day_From_Start >= 120 & Day_From_Start < 180 ~ 4,
      Day_From_Start >= 180 & Day_From_Start < 270 ~ 5,
      Day_From_Start >= 270 & Day_From_Start < 365~ 6,
      Day_From_Start >= 365 & Day_From_Start < 545~ 7,
      Day_From_Start >= 545 & Day_From_Start < 730~ 8,
      Day_From_Start >= 730 ~ 9
    )
  )



ICI_period <- ICI_period %>%
  filter(gap_over_60 == TRUE)

ICI_period <- ICI_period %>%
  distinct(PERSON_ID, ICI_prescrib_TDPS, .keep_all = TRUE)

patient_mono_time <- left_join(patient_mono_time, select(ICI_period, PERSON_ID, ICI_prescrib_TDPS, gap_over_60), by = c('PERSON_ID', 'ICI_prescrib_TDPS'))

patient_mono_time <- patient_mono_time %>%
  mutate(
    gap_over_60 = if_else(is.na(gap_over_60), 0, as.numeric(gap_over_60))
  )





patient_mono_time <- patient_mono_time %>%
  mutate(
    timeseries_index = case_when(
      ICI_prescrib_TDPS == 1 ~ COHORT_START_DATE,
      ICI_prescrib_TDPS == 2 ~ COHORT_START_DATE + 30,
      ICI_prescrib_TDPS == 3 ~ COHORT_START_DATE + 60,
      ICI_prescrib_TDPS == 4 ~ COHORT_START_DATE + 120,
      ICI_prescrib_TDPS == 5 ~ COHORT_START_DATE + 180,
      ICI_prescrib_TDPS == 6 ~ COHORT_START_DATE + 270,
      ICI_prescrib_TDPS == 7 ~ COHORT_START_DATE + 365,
      ICI_prescrib_TDPS == 8 ~ COHORT_START_DATE + 545,
      ICI_prescrib_TDPS == 9 ~ COHORT_START_DATE + 730,
    ))





# irAE covariates 
irAE_total <- left_join(irAE_total, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')
irAE <- irAE_total %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_irAE_days = as.numeric(difftime(CONDITION_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()
irAE <- irAE %>%
  mutate(
    irAE_occur = case_when(
      Index_irAE_days >= -365 & Index_irAE_days < 0 ~ 1,
      Index_irAE_days >= 0 & Index_irAE_days < 30 ~ 2,
      Index_irAE_days >= 30 & Index_irAE_days < 60 ~ 3,
      Index_irAE_days >= 60 & Index_irAE_days < 120 ~ 4,
      Index_irAE_days >= 120 & Index_irAE_days < 180 ~ 5,
      Index_irAE_days >= 180 & Index_irAE_days < 270 ~ 6,
      Index_irAE_days >= 270 & Index_irAE_days < 365 ~ 7,
      Index_irAE_days >= 365 & Index_irAE_days < 545 ~ 8,
      Index_irAE_days >= 545 & Index_irAE_days < 730 ~ 9
    )
  )
irAE <- irAE %>%
  group_by(PERSON_ID, irAE_occur) %>%
  filter(CONDITION_START_DATE == max(CONDITION_START_DATE)) %>%
  ungroup()
irAE <- irAE %>%
  group_by(PERSON_ID) %>%
  filter(irAE_occur == min(irAE_occur)) %>%
  ungroup()
irAE <- irAE %>%
  distinct(PERSON_ID, .keep_all = TRUE)
irAE <- na.omit(irAE, subset = "irAE_occur")
patient_mono_time <- left_join(patient_mono_time, select(irAE, c('PERSON_ID', 'irAE_occur')), by = 'PERSON_ID')

patient_mono_time <- patient_mono_time %>%
  mutate(irAE = if_else(is.na(irAE_occur),
                          0,
                          if_else(irAE_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time <- select(patient_mono_time, -irAE_occur)



##Hospitalization within 15 days prior
admission_raw <- read.csv("addmission.CSV")[,-1]
ER_raw <- read.csv("ER.CSV")[,-1]

colnames(admission_raw)
colnames(ER_raw)

admission <- select(admission_raw, PERSON_ID, VISIT_START_DATE)
ER <- select(ER_raw, PERSON_ID, VISIT_START_DATE)

hospital <- rbind(admission, ER)
hospital <- unique(hospital)


hospital <- left_join(hospital, select(patient, c(PERSON_ID, COHORT_START_DATE)), by = 'PERSON_ID')

hospital <- hospital %>%
  group_by(PERSON_ID) %>%
    mutate(
      COHORT_START_DATE = as.Date(COHORT_START_DATE)
    ) %>%
    mutate(
      Index_hosp_days = as.numeric(difftime(VISIT_START_DATE, COHORT_START_DATE, units = "days")+1)
    ) %>%
    ungroup()

hospital <- hospital %>%
  mutate(
    hospital_occur = case_when(
      Index_hosp_days > -15 & Index_hosp_days <= 0 ~ 1,
      Index_hosp_days > 15 & Index_hosp_days <= 30 ~ 2,
      Index_hosp_days > 45 & Index_hosp_days <= 60 ~ 3,
      Index_hosp_days > 105 & Index_hosp_days <= 120 ~ 4,
      Index_hosp_days > 165 & Index_hosp_days <= 180 ~ 5,
      Index_hosp_days > 255 & Index_hosp_days <= 270 ~ 6,
      Index_hosp_days > 350 & Index_hosp_days <= 365 ~ 7,
      Index_hosp_days > 530 & Index_hosp_days <= 545 ~ 8,
      Index_hosp_days > 715 & Index_hosp_days <= 730 ~ 9
    )
  )

hospital <- hospital %>%
  group_by(PERSON_ID, hospital_occur) %>%
  filter(VISIT_START_DATE == max(VISIT_START_DATE)) %>%
  ungroup()

hospital$ICI_prescrib_TDPS <- hospital$hospital_occur

hospital <- hospital %>%
  distinct(PERSON_ID, hospital_occur, .keep_all = TRUE)

hospital <- na.omit(hospital, subset = "hospital_occur")


patient_mono_time <- left_join(patient_mono_time, select(hospital, c('PERSON_ID', 'hospital_occur','ICI_prescrib_TDPS')), by = c('PERSON_ID','ICI_prescrib_TDPS'))
patient_mono_time <- patient_mono_time %>%
  mutate(hospital_sever = if_else(is.na(hospital_occur), 
                          0, 
                          1))
patient_mono_time <- select(patient_mono_time, -hospital_occur)




write.csv(patient_mono_time, file = "patient_mono_time_sens_sever_250318.CSV")



```





Time-dependent PS matching

```{r}



# Exclude variables with >30% missingness
missing_percentages <- colSums(is.na(patient_mono_time)) / nrow(patient_mono_time) * 100
print(missing_percentages)


#imputation
patient_mono_time <- patient_mono_time %>%
  group_by(PERSON_ID) %>%
  mutate(ANC_lab = ifelse(is.na(ANC_lab), mean(ANC_lab, na.rm = TRUE), ANC_lab),
         NLR_cal = ifelse(is.na(NLR_cal), mean(NLR_cal, na.rm = TRUE), NLR_cal)) %>%
  ungroup()

## weight imputation
patient_mono_time <- patient_mono_time %>%
  group_by(PERSON_ID) %>%
  mutate(weight = ifelse(is.na(weight), mean(weight, na.rm = TRUE), weight)) %>%
  ungroup()




patient_mono_time <- patient_mono_time %>%
  mutate(ANC_category = case_when(
    ANC_lab < 1500 ~ 1,
    ANC_lab >= 1500 & ANC_lab <= 8000 ~ 2,
    ANC_lab > 8000 ~ 3
  ))

patient_mono_time <- patient_mono_time %>%
  mutate(NLR_category = case_when(
    NLR_cal < 3 ~ 1,
    NLR_cal >= 3 & NLR_cal <= 5 ~ 2,
    NLR_cal > 5 ~ 3
  ))



patient_mono_time_8 <- patient_mono_time %>% filter(is.na(ad_steroid) | ad_steroid != 9)
print(TARcheck <- patient_mono_time_8 %>% filter(TARGET==1) %>% summarise(count = n_distinct(PERSON_ID)))



library("tableone")


colnames(patient_mono_time_8)


patient_mono_time_ps <- select(patient_mono_time_8, PERSON_ID, ICI_prescrib_TDPS, ad_steroid, age, sex, squamouslc, durvalumab_use, ANC_category, NLR_category, AUTOIMMUNE, LIVER_INSUFFICIENCY, RENAL_INSUFFICIENCY, COPD, metastasis, RT_GKS, pre_chemo, pre_TKI_ALK, pre_steroid, ANTI_BIOTICS, NSAID, PPI, ACEI, OPIOID, irAE, Sepsis, hospital_sever)


library("rsmatch")
library("survival")
library("nbpMatching")



str(patient_mono_time_ps)
missing_percentages <- colSums(is.na(patient_mono_time_ps)) / nrow(patient_mono_time_ps) * 100
print(missing_percentages)



patient_mono_time_ps <- patient_mono_time_ps %>%
  mutate(sex = as.factor(sex))
colnames(patient_mono_time_ps)




patient_mono_time_pair <- coxpsmatch(
  n_pairs = 10000,
  data = patient_mono_time_ps,
  id = "PERSON_ID",
  time = "ICI_prescrib_TDPS",
  trt_time = "ad_steroid",
  covariates = c("metastasis", "RT_GKS", "age", "sex", "ANC_category", "NLR_category", "AUTOIMMUNE", "LIVER_INSUFFICIENCY", "RENAL_INSUFFICIENCY", "COPD", "pre_chemo", "pre_TKI_ALK", "pre_steroid", "ANTI_BIOTICS", "NSAID", "PPI", "ACEI", "OPIOID", "irAE", "squamouslc", "durvalumab_use", "Sepsis", "hospital_sever"),
  exact_match = NULL,
  options = list(time_lag = FALSE)
)



patient_mono_time_pair_filt <- patient_mono_time_pair %>%
  filter(!is.na(type))

patient_mono_time_pair_filt <- patient_mono_time_pair_filt %>%
  arrange(pair_id)

patient_mono_time_ps <- left_join(patient_mono_time_pair_filt, patient_mono_time, by ="PERSON_ID")





# index date matching
patient_mono_time_ps <- patient_mono_time_ps %>%
  group_by(pair_id) %>%
  mutate(
    index = if (any(type == "trt")) first(ad_steroid[type == "trt"], default = NA) else NA,
    index_fu = case_when(
      type == "trt" & ICI_prescrib_TDPS == index ~ as.character(timeseries_index),
      type == "all" & ICI_prescrib_TDPS == index ~ as.character(timeseries_index),
      type == "all" & !any(ICI_prescrib_TDPS == index) ~ "Exclusion",
      TRUE ~ NA_character_
    )
  ) %>%
  ungroup() %>%
  select(-index)


# pair-wise exclusion
#  1) Drop records where the index_date cannot be matched
exclusion_rows <- patient_mono_time_ps %>%
  filter(index_fu == "Exclusion")
print(exclusion_rows)


check <- patient_mono_time_ps %>%
  filter((type == "all" & TARGET != 0) | (type == "trt" & TARGET != 1))
print(check)


# 2) Exclude participants who are not on ICI monotherapy at the new index date
invalid_pair_ids <- patient_mono_time_ps %>%
  filter(other_CA_START_DATE <= index_fu) %>%
  pull(pair_id)

print(invalid_pair_ids)


patient_mono_time_ps_pex <- patient_mono_time_ps %>%
  filter(!pair_id %in% invalid_pair_ids)


unique(patient_mono_time_ps_pex$PERSON_ID)


write.csv(patient_mono_time_ps, file = "patient_mono_time_ps_sens_sever_250318.CSV")
write.csv(patient_mono_time_ps_pex, file = "patient_mono_time_pex_sens_sever_250318.CSV")






library("tableone")


#---------------------------------------------Convert to cumulative variables: Dyspnea, Radiotherapy, Whole_Brain_RT, RTsurgery, gap_over_60


patient_mono_time_ps_demo <- patient_mono_time_ps_pex



colnames(patient_mono_time_ps_demo)


# Dyspnea
dyspnea <- dyspnea %>%
  group_by(PERSON_ID) %>%
  filter(dyspnea_occur == min(dyspnea_occur)) %>%
  ungroup()

dyspnea <- dyspnea %>%
  distinct(PERSON_ID, .keep_all = TRUE)

dyspnea <- na.omit(dyspnea, subset = "dyspnea_occur")

patient_mono_time_ps_demo <- left_join(patient_mono_time_ps_demo, select(dyspnea, c('PERSON_ID', 'dyspnea_occur')), by = 'PERSON_ID')
patient_mono_time_ps_demo <- patient_mono_time_ps_demo %>%
  mutate(Dyspnea = if_else(is.na(dyspnea_occur), 
                          0, 
                          if_else(dyspnea_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time_ps_demo <- select(patient_mono_time_ps_demo, -dyspnea_occur)
patient_mono_time_ps_demo <- unique(patient_mono_time_ps_demo)


# RT_GKS
RT_GKS <- RT_GKS %>%
  group_by(PERSON_ID) %>%
  filter(RT_occur == min(RT_occur)) %>%
  ungroup()

RT_GKS <- RT_GKS %>%
  distinct(PERSON_ID, .keep_all = TRUE)

RT_GKS <- na.omit(RT_GKS, subset = "RT_occur")

patient_mono_time_ps_demo <- left_join(patient_mono_time_ps_demo, select(RT_GKS, c('PERSON_ID', 'RT_occur')), by = 'PERSON_ID')
patient_mono_time_ps_demo <- patient_mono_time_ps_demo %>%
  mutate(RT_GKS = if_else(is.na(RT_occur), 
                          0, 
                          if_else(RT_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time_ps_demo <- select(patient_mono_time_ps_demo, -RT_occur)




# gap_over_60

ICI_period$gap_occur <- ICI_period$ICI_prescrib_TDPS

ICI_period <- ICI_period %>%
  group_by(PERSON_ID) %>%
  filter(gap_occur == min(gap_occur)) %>%
  ungroup()

ICI_period <- ICI_period %>%
  distinct(PERSON_ID, .keep_all = TRUE)

ICI_period <- na.omit(ICI_period, subset = "gap_occur")

patient_mono_time_ps_demo <- left_join(patient_mono_time_ps_demo, select(ICI_period, c('PERSON_ID', 'gap_occur')), by = 'PERSON_ID')
patient_mono_time_ps_demo <- patient_mono_time_ps_demo %>%
  mutate(gap_over_60 = if_else(is.na(gap_occur), 
                          0, 
                          if_else(gap_occur <= ICI_prescrib_TDPS, 1, 0)))
patient_mono_time_ps_demo <- select(patient_mono_time_ps_demo, -gap_occur)

patient_mono_time_ps_demo <- unique(patient_mono_time_ps_demo)



patient_mono_time_ps_demo <- patient_mono_time_ps_demo %>%
  filter(!is.na(index_fu))

colnames(patient_mono_time_ps_demo)

patient_mono_time_ps_demo <- unique(patient_mono_time_ps_demo)

str(patient_mono_time_ps_demo)


write.csv(patient_mono_time_ps_demo, file = "patient_mono_time_ps_demo_sens_sever_250318.CSV")


patient_mono_time_ps_demo <- select(patient_mono_time_ps_demo, -c("index_fu", "COHORT_START_DATE", "COHORT_END_DATE", "Steroid_index_date", "other_CA_START_DATE", "timeseries_index"))

factorVars <- colnames(select(patient_mono_time_ps_demo, other_CA, pembrolizumab_use, nivolumab_use, atezolizumab_use, durvalumab_use, ICI_dupl, squamouslc, sex, AUTOIMMUNE, LIVER_INSUFFICIENCY, RENAL_INSUFFICIENCY, COPD, ANTI_BIOTICS, NSAID, PPI, ACEI, OPIOID, pre_steroid, pre_chemo, pre_TKI_ALK, ICI_prescrib_TDPS, ad_steroid, Sepsis, Pneumonitis, Dyspnea, RT_GKS, Brainmeta, Bonemeta, Livermeta, metastasis, gap_over_60, ANC_category, NLR_category, irAE, hospital_sever))

table2 <- CreateTableOne(factorVars = factorVars,
                         data = patient_mono_time_ps_demo,
                         strata = 'TARGET',)


table2 <- as.data.frame(print(table2, quote = FALSE, noSpaces = TRUE))
write.csv(table2, file = "table2_sens_sever.CSV")



```




